The Parareal Algorithm for Parallel Solution of ODEs

### 2014-12-09 poster/main.tex
more poster files

We implement the parareal algorithm to solve various ODEs and compare in
terms of accuracy, speedup and efficiency. We explore the effect of using
different solution operators (a coarse and a fine as required by the method).
Our code is written in python using mpi4py to parallelize. We observe
light speedups in our small scale tests up to 64 processors on Harvard's
Odyssey cluster. We analyze possible reasons as to why our implementation
may not be ideal and some tradeoffs that are taken in the parareal
implementation. Further optimizations are proposed and rationalized as next
steps including ports of to C++ and BLAS libraries which is a more
performance-oriented, lower-level language than Python. Data is shown for
solving for various complexities of ODEs from the basic exponential to
modeling sound wave propagation as we did in homework 4.

@@

The parareal algorithm, developed by Lions, Maday and Turinici in 2001, is
a general algorithm that allows parallelization in time.sIt does so by using a
cheaper (lower order, lesser resolution) approximation first, and then makes
corrections in parallel. The entire process is then iterative and tuned for a
given amount of iterations. Therefore, the algorithm is $k$ repetitions of a
serial coarse method updated by a finer method run in parallel for subsections.

@@

\emph{Theorem}: The order of accuracy of the parareal method with coarse
solution operator $g_{\Delta t}$ and fine operator $g$ is $mk$. Can be proved by
induction.

@@

With the parareal method, it is possible to combine ODE solvers. Stability
region depends on both $g_{\Delta t}$ and $g_{\textnormal{fine}}$, and the
equation being solved.

@@

Let's say the coarse operator runs in time $t$, and the fine operator runs in
time $Q \cdot t$. Assume we have $N$ processors and we perform $k$ correction
steps. Then, the runtime of parareal is, assuming negligible setup and
aggregation time,

@@

We tested some stuff on Odyssey.

>>>>

We implement the Parareal algorithm to solve various ODEs and compare in
terms of accuracy, speedup and efficiency. We explore the effect of using
different solution operators (a coarse and a fine as required by the method).
Our code is written in python using mpi4py to parallelize. We observe
light speedups in our small scale tests up to 64 processors on Harvard's
Odyssey cluster. We analyze possible reasons as to why our implementation
may not be ideal and some tradeoffs that are taken in the Parareal
implementation. Further optimizations are proposed and rationalized as next
steps including ports of to C++ and BLAS libraries which is a more
performance-oriented, lower-level language than Python. Data is shown for
solving for various complexities of ODEs from the basic exponential to
modeling sound wave propagation as we did in homework 4.

@@

The Parareal algorithm, developed by Lions, Maday and Turinici in 2001, is
a general algorithm that allows parallelization in time.sIt does so by using a
cheaper (lower order, lesser resolution) approximation first, and then makes
corrections in parallel. The entire process is then iterative and tuned for a
given amount of iterations. Therefore, the algorithm is $k$ repetitions of a
serial coarse method updated by a finer method run in parallel for subsections.

@@

\emph{Theorem}: The order of accuracy of the Parareal method with coarse
solution operator $g_{\Delta t}$ and fine operator $g$ is $mk$. Can be proved by
induction.

@@

With the Parareal method, it is possible to combine ODE solvers. Stability
region depends on both $g_{\Delta t}$ and $g_{\textnormal{fine}}$, and the
equation being solved.

@@

Let's say the coarse operator runs in time $t$, and the fine operator runs in
time $Q \cdot t$. Assume we have $N$ processors and we perform $k$ correction
steps. Then, the runtime of Parareal is, assuming negligible setup and
aggregation time,

@@

We implemented the Parareal algorithm in Python, using mpi4py to parallelize it.
Our coarse operator was forward Euler step, with 100 steps, while the fine
operator was forward Euler, with $Q * 100$ steps, where $Q$ is the quality
factor. We tested the Parareal algorithm on two sets of differential equations:

\[
y'(t) = f(t, y) = \lambda y, \, y(0) = 1
\]
\[
y''(t) + 2y'(t) + 5y(t) = 0, \, y(0) = 1 \, y'(0) = 0.
\]

We ran the algorithm on different numbers of processors and varied $k$ and $Q$.
